Effects of below-ground sink limitation on growth and carbon balance of Eucalyptus seedlings

### 2015-07-23 manuscript/manuscript_pve.Rmd
minor edits to manuscript

Interpreting limitations to plant growth requires understanding of the balance between carbon (C) source and sink activity in order to assess C allocation and biomass partitioning. This study used manipulations of soil volume to test how growth is coupled to physiology, allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally sown seedlings (free). Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that soil volume effect would be largest in the smallest containers, negatively impacting mass partitioning belowground. The accumulation of leaf non-structural carbohydrates, resulting from reduced sink strengh, was expected to correleltae to reductions in photosynthetic capacity. We observed a negative container effect on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes mass, partitioning to leaves, stems and roots was conserved after 120 days. Photosynthetic capacity was also significantly reduced in containers, and was related to both leaf nitrogen content and starch accumulation. We developed a seedling growth model that utilized leaf photosynthesis (A) rates to allocate daily C uptake towards mass growth of stems, leaves and roots. We then asked whether the observed reductions in A explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in the down regulation of A, these reductions were not significant enough to explain observed growth responses. Thus, as A and growth were not coordinated an excess pool of non-biomass C resulted in seedlings with soil volume restriction. This research highlights the need to further utilize mass balance approaches when evaluating plant C allocation and confirms that A and growth are not always directly related.

@@

\
It was further necessary to calculate a self-shading parameter ($\sigma$~s~) when scaling leaf A with total plant leaf area. This was accomplished by utilizing 61 previously digitized Eucalyptus seedlings, covering 5 total species which include *E. tereticornis*, from Duursma et al. [-@duursma2012light] to run in 'YplantQMC' package in R (fix cite YplantQMC) to build a 3D plant structure based on digitized metrics of plant allometry and crown structure. Inputting the same treatment specific physiological parameters listed above, 'YplantQMC' outputs total A, using total leaf area, for seedlings assuming self-shading as well as for a full sun large horizontal leaf.  The ratio of total A with self-shading to horizontal leaf was then used to calculate $\sigma$~s~ for each of the 61 digitized seedling, independently for each treatment. Next, the linear relationship between $\sigma$~s~ and total leaf area was for determined across digitized seedlings, within each treatment. For the growth model, $\sigma$~s~ was then predicted for each daily time step using the previous days cumulative leaf area and this value  was then applied to C~day,i~. All default parameters used in model simulations are reported in Table.3.

@@

## Growth and morphology metrics
In this field study, colder temperatures and reductions in cumulative PPFD per day (Fig. `r airvarsfig`) most likely lead to the reduced growth in the free seedlings in the final weeks of the experiment (Fig. `r allomfig`).  Combined with severe growth reductions in the smallest container volumes the experiment was chosen to be harvested after 120 days. Over this duration height, diameter, and leaf area diverged between container volumes (Fig. `r allomfig`).  First, seedling leaf area significantly diverged between soil volumes (P < 0.026) during the 5^th^ week of the experiment. Following this period both height (8^th^} week) and then diameter (9^th^ week) significantly deviated across soil volumes (P < 0.002 & 0.001, respectively).  Negative growth effects then manifested as severely reduced height gain and declining leaf area through time with small soil volumes across the final two months of the experiment. Seedlings maintained diameter growth throughout the experiment, although marginal with smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (P < 0.001).  Increases in both final stem diameter (P < 0.001) and cumulative leaf area (both P < 0.001) were found with increasing soil volume and these differences were driven mainly by the largest container and the free seedling.

\
Total seedling biomass at harvest was significantly different across container volumes (P < 0.001) and with free seedlings (P < 0.001, Table. 1). We analyzed the relationship between biomass growth with each fold increase in soil volume and found an increase of 34 % with a doubling of pot size, consistent with the meta-analysis of Poorter et al. [-@poorter2012pot]. Additionally, plant biomass was highly correlated with total leaf area across all treatments (R^2^ = 0.97, P < 0.001). Differences in biomass partitioning to leaves, stems, and roots were not different across soil volumes after variation in seedling biomass within treatments was factored in the analysis (Fig.`r allocfig`a,b). Across all treatments, the final harvested root:shoot was conserved in these seedlings, with a slightly higher shoot than root mass on average ($\overline{x}$ = 0.904, Fig. `r allomfig`c).

\
Overall, SRL was higher in seedlings in containers compared to free seedlings but not significantly in every soil volume treatment   (Table. 1). Over the duration of the experiment SLA was higher in free seedlings but was not different across containers sizes (Table. 1, P < 0.001) and this pattern was evident in the first gas exchange measurement campaign (P < 0.001).

## Leaf chemistry
Leaf N % was significantly higher in free seedlings and the largest container volume at the onset of gas exchange measurements (6th week, P < 0.001).  Over the remaining duration of the experiment the smallest container volume had a significant reduction in leaf N % compared to other soil volumes, while free seedlings maintained the highest leaf N % (Table. 1, P < 0.001).  Leaf starch content in the smallest container was double that of free seedlings (P=0.039), while leaf soluble sugars did not differ across treatments throughout the experiment (Table. 1).  Differences in leaf starch between the free seedling and the smallest container were also evident during the first gas exchange campaign (P = 0.001).

## Gas exchange and photosynthetic parameters
A~sat~ and A~max~ were both significantly higher in the largest container volume and the free seedling at the first measurement campaign (both P < 0.001). Across all measurement campaigns A~sat~ (Fig. `r asatfig`) and A~max~ (Table 2.) were consistently higher in free seedlings than in containers (both P < 0.001). The interaction between photosynthetic capacity, leaf starch, and leaf N on a mass basis was marginally significant (P = 0.058) but A~max~ was highly correlated to both leaf N content and leaf starch (both P < 0.001). Across all measurement campaign A~max~ was higher when foliar N was also higher, usually associated with low levels of leaf starch (Fig. `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Fig. `r achemfig`b)

\
The photosynthetic parameters J~max~ and Vc~max~ were not different within treatments at the beginning and end of gas exchange campaigns, therefore the parameter means per treatment are reported here (Table. 2).  Overall, both J~max~ and Vc~max~ were significantly higher in free seedlings with little variation between soil volume treatments (P = 0.001 & 0.002, respectively). Leaf dark respiration rates were not significantly different across soil volumes (Table. 2). The g~1~ parameter, generated for each seedling from the Medlyn et al [-@medlyn2012reconciling] optimal stomatal conductance model, was lowest in the free seedling and was marginally different across soil volume treatments (Table. 2). Predicted values of g~s~, using the g~1~ parameter, where highly correlated with observed values (R^2^ = .74, P < 0.001, data not shown).

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 mPa across all seedlings, respectively. Although g~s~ in free seedlings was genearlly higher htan those in containers (Table 2. P=0.001), the mean rates for all seedlings were high at 0.37 mol H~2~0 m^-2^ s^-1^ and did not decline significantly across the experiment duration. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table. 1). Combined these indices provide strong evidence that water stress was not apparent on these well-watered seedlings throughout the experiment. Soil N % at harvest was not different across soil volumes ($\bar{x}$ = 04.5 %) and decreased approximately 3 % across all containers over the experiment duration. This indicates that nutrient leaching from free seedlings or from draining of containers following natural rainfall events did not differ between treatments.

## Modelling seedling biomass
Model M~0~, was able to converge on an optimum LMF (21.6 %) which predicted the mean harvest total biomass of free seedlings within 1.2 %. Using this optimized LMF, the total biomass of modeled seedlings with soil volume restriction (M~1~) was on average 23&pm;2.4 g C more than measured seedlings when comparing against predicted total net leaf C gain (Fig. `r massmodelfig`a). Thus, seedling C mass was overestimated by an average of 50&pm;8.7 % in modeled seedlings across the reductions in daily C assimilation per treatment (Fig. `r massmodelfig`b). As a result, the reductions in leaf A were not sufficient enough to explain the reduction in harvested seedling biomass with soil volume restriction. This resulted in an unexplained pool of excess C, generated from A, which was not allocated to biomass.

\
The remaining model simulations tested possible C allocation scenarios to account for this excess pool of C. Constraining the model by treatment-specific LMF from the final harvest (M~2~) still overestimated seedling total C by 32&pm;11.1 %, but provided the most improved model predictions (Fig. `r massmodelfig2`a). Using harvested LMF, however, does not capture the increase in senescence of seedlings in small containers (Fig. `r allomfig`c). Thus, the 7-61 % reduction in measured harvested leaf mass fraction from large to small soil volumes compared to the optimized LMF (M~0~) represents an underestimation of realized leaf C allocation. Increases of 50 % in non-leaf tissue respiration (M~3~) improved biomass estimates slightly but overestimated mass C by an average of 46&pm;9.3 % in seedlings with soil volume restriction (Fig. `r massmodelfig2`b). It was further determined that non-tissue respiration rates in modeled seedlings would need to increase by ca. 250 % to account for this entire pool of C.

@@

\
With high rates of g~s~, non-limiting leaf water potential and consistent leaf $\delta$^13^C across soil volume treatments there was little evidence of water stress inducing a limitation to A. This finding is consistent with other container size studies without drought treatments.  For example, reduced A~max~ in cotton seedlings grown at elevated CO~2~ was attributed to sink-limited feedback inhibition from inadequate rooting volume, not decreased g~s~ [@thomas1991root]. Additionally, severe reductions in A in coffee plants was not attributed to impacts of container size on leaf water potentials or g~s~ [@ronchi2006growth]. Here, it is likely that reductions in A of these well watered seedlings was instead the result of limiting soil nutrients on belowground sink strength.

\
Although the soil N pool declined evenly across all treatments leaf N was lowest in the smallest containers, suggesting that sink limitation was the greatest in these containers. This makes sense as small containers may reduce N uptake, either from physical root restriction or decreased supply, which affects growth, Rubisco limitation, sugar metabolism, and carbohydrate partitioning between source and sink tissues [@stitt1991rising; @hermans2006plants]. Mycorrhizal colonization could also have been affected in containers. Whether unrestricted mycorrhizal recruitment facilitated the increase in leaf N uptake in free seedlings is unknown. Regardless of whether sink limitation occurred from root or mycorrhizal restriction, the stark contrast in leaf N between free seedlings and seedlings in containers was significantly correlated to reductions in A.

\
As both rooting space and resources were finite in containers, the inability of seedlings to increase the capacity of the belowground C sink resulted in the buildup of C assimilate in leaves. The feedback inhibition of A from starch accumulation has been proposed, yet it is still not known whether there is a starch threshold that triggers the down-regulation process (Nebauer et al. 2011).  Here, declines in A~max~ where correlated with higher starch content throughout the experiment.  This agrees with findings from Equiza et a. (2006) where photosynthetic downregulation from reduced sink strength showed a greater correlation with starch than with sugar content in a deciduous conifer.  As starch content in the smallest containers was nearly double that of free seedlings this suggests the response of A to sink inhibition was regulated by this accumulation, as hypothesized.

@@

\
As partitioning to fine roots did not change this provides evidence against an optimal foraging strategy for seedlings in containers. This could be because lateral root development is affected by inanimate obstacles and avoiding growth towards container walls could improve the efficiency of resource allocation [@falik2005root]. Also, the sensitivity of roots to their own exudates near obstructions may be used to adjust growth before stressful conditions occur [@semchenko2008foraging]. Alternatively, physical restriction of root proliferation could have impacted root development or morphology prior to shifts in mass partitioning. Here, increases in root length where detected in several of the soil volume treatments . This is not surprising, as plants in containers have been shown to have different root morphology to field grown plants as roots essentially compete with themselves for nutrients [@nesmith1998effect]. Poor soil quality and root restriction, however, likely decreased the capacity of this morphological response to increase N uptake.

@@

\
Increasing rates of non-leaf respiration (M~3~) improved biomass predictions but to a far lesser extent than changes to leaf C allocation. This finding is still noteworthy as it shows that that sink limitation could differentially affect respiration of component tissues. Additionally, keeping rates of respiration fixed in growth models may underestimate impacts of environmental change on tissue respiration above and belowground. The fraction of photosynthate used in respiration varies substantially among species and environments  and is sensitive to changes in growth rates [@Lambers2008]. As C balance is a delicate equilibrium between fluxes of A and respiration, partial accounting of C dynamics can easily lead to erroneous conclusions (Valentini et al. 2000).We agree with Delucia et al. [-@delucia2007forest] that it is likely inappropriate to assume that respiration is a constant fraction of gross primary production in models. Our findings reveal that a combination of different mechanisms, beyond A, is likely at play in driving the observed seedling biomass response to sink manipulation. However, the degree to which these mechanisms will regulate a growth will undoubtedly shift across different experimental manipulations and plant species.

## Conclusions
It has long been know that conditions which affect the photosynthetic process affect growth but that there is also an interrelationship between growth and A [@sweet1966role]. Here, the fate of assimilated C available for plant growth varied between naturally sown seedlings and seedlings with belowground sink inhibition. First, this is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles but how these results actually reflect field-grown plants has seldom been studied. Second, our finding add more evidence that A and growth are not always entirely synced, an important distinction often missed in studies that manipulate source/sink activity. Korner [-@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control A through the demand for C. Although this may be true, we argue that attempts to quantify or at least predict the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents.  A lack of knowledge regarding C allocation restricts our ability to achieve mass balance and is a major obstacle in understanding the coordination between A and growth. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

@@

**Table 1**. Responses of plant and leaf characterisitics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Seedling mass, SRL, root nitrogen and leaf $\delta$^13C values are from final harvest. Values of leaf starch, sugars, nitrogen and SLA represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

\
**Table 2**. Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Units for A~max~ and R~dark~ are $\mu$mol m^-2^ s^-1^ and g~s~ are mol m^-2^ s^-1^, each at at 25 &deg;C. Values of A~max~, g~s~ and g~1~ represent overall means across measurement campaigns (n = 6). R~dark~, J~max~ and Vc~max~ values are means of two measurement campaigns at beginning and end of gas exchange measurements. Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

>>>>

Interpreting limitations to plant growth requires understanding of the balance between carbon (C) source and sink activity in order to assess C allocation and biomass partitioning. This study used manipulations of soil volume to test how growth is coupled to physiology, allocation, and sink activity in *Eucalyptus tereticornis* seedlings. We grew seedlings in a large range of container sizes and planted containers flush to the soil alongside naturally sown seedlings (free). Reduced soil volume was expected to induce rapid negative effects on growth and physiology compared to free seedlings. It was hypothesized that soil volume effect would be largest in the smallest containers, negatively impacting mass partitioning belowground. The accumulation of leaf non-structural carbohydrates, resulting from reduced sink strength, was expected to correlate to reductions in photosynthetic capacity. We observed a negative container effect on aboveground growth soon after the experiment started. Although growth was consistently different across soil volumes mass, partitioning to leaves, stems and roots was conserved after 120 days. Photosynthetic capacity was also significantly reduced in containers, and was related to both leaf nitrogen content and starch accumulation. We developed a seedling growth model that utilized leaf photosynthesis (A) rates to allocate daily C uptake towards mass growth of stems, leaves and roots. We then asked whether the observed reductions in A explained the observed differences in seedling biomass. We found that although belowground sink limitation resulted in the down regulation of A, these reductions were not significant enough to explain observed growth responses. Thus, as A and growth were not coordinated an excess pool of non-biomass C resulted in seedlings with soil volume restriction. This research highlights the need to further utilize mass balance approaches when evaluating plant C allocation and confirms that A and growth are not always directly related.

@@

\
It was further necessary to calculate a self-shading parameter ($\sigma$~s~) when scaling leaf A with total plant leaf area. This was accomplished by utilizing 61 previously digitized Eucalyptus seedlings, covering 5 total species which include *E. tereticornis*, from Duursma et al. [-@duursma2012light] to run in 'YplantQMC' package in R (fix cite YplantQMC) to build a 3D plant structure based on digitized metrics of plant allometry and crown structure. Inputting the same treatment specific physiological parameters listed above, 'YplantQMC' outputs total A, using total leaf area, for seedlings assuming self-shading as well as for a full sun large horizontal leaf.  The ratio of total A with self-shading to horizontal leaf was then used to calculate $\sigma$~s~ for each of the 61 digitized seedling, independently for each treatment. Next, the linear relationship between $\sigma$~s~ and total leaf area was for determined across digitized seedlings, within each treatment. For the growth model, $\sigma$~s~ was then predicted for each daily time step using the previous days cumulative leaf area and this value  was then applied to C~day,i~. All default parameters used in model simulations are reported in Table S1.

@@

## Growth and morphology metrics
In this field study, colder temperatures and reductions in cumulative PPFD per day (Figure `r airvarsfig`) most likely lead to the reduced growth in the free seedlings in the final weeks of the experiment (Figure `r allomfig`).  Combined with severe growth reductions in the smallest container volumes the experiment was chosen to be harvested after 120 days. Over this duration height, diameter, and leaf area diverged between container volumes (Figure `r allomfig`).  First, seedling leaf area significantly diverged between soil volumes (P < 0.026) during the 5^th^ week of the experiment. Following this period both height (8^th^} week) and then diameter (9^th^ week) significantly deviated across soil volumes (P < 0.002 & 0.001, respectively).  Negative growth effects then manifested as severely reduced height gain and declining leaf area through time with small soil volumes across the final two months of the experiment. Seedlings maintained diameter growth throughout the experiment, although marginal with smaller soil volumes in the final month. Final seedling height significantly increased with increasing soil volume (P < 0.001).  Increases in both final stem diameter (P < 0.001) and cumulative leaf area (both P < 0.001) were found with increasing soil volume and these differences were driven mainly by the largest container and the free seedling.

\
Total seedling biomass at harvest was significantly different across container volumes (P < 0.001) and with free seedlings (P < 0.001, Table 1). We analyzed the relationship between biomass growth with each fold increase in soil volume and found an increase of 34 % with a doubling of pot size, consistent with the meta-analysis of Poorter et al. [-@poorter2012pot]. Additionally, plant biomass was highly correlated with total leaf area across all treatments (R^2^ = 0.97, P < 0.001). Differences in biomass partitioning to leaves, stems, and roots were not different across soil volumes after variation in seedling biomass within treatments was factored in the analysis (Figure`r allocfig`a,b). Across all treatments, the final harvested root:shoot was conserved in these seedlings, with a slightly higher shoot than root mass on average ($\overline{x}$ = 0.904, Figure `r allomfig`c).

\
Overall, SRL was higher in seedlings in containers compared to free seedlings but not significantly in every soil volume treatment   (Table 1). Over the duration of the experiment SLA was higher in free seedlings but was not different across containers sizes (Table 1, P < 0.001) and this pattern was evident in the first gas exchange measurement campaign (P < 0.001).

## Leaf chemistry
Leaf N % was significantly higher in free seedlings and the largest container volume at the onset of gas exchange measurements (6th week, P < 0.001).  Over the remaining duration of the experiment the smallest container volume had a significant reduction in leaf N % compared to other soil volumes, while free seedlings maintained the highest leaf N % (Table 1, P < 0.001).  Leaf starch content in the smallest container was double that of free seedlings (P=0.039), while leaf soluble sugars did not differ across treatments throughout the experiment (Table 1).  Differences in leaf starch between the free seedling and the smallest container were also evident during the first gas exchange campaign (P = 0.001).

## Gas exchange and photosynthetic parameters
A~sat~ and A~max~ were both significantly higher in the largest container volume and the free seedling at the first measurement campaign (both P < 0.001). Across all measurement campaigns A~sat~ (Figure `r asatfig`) and A~max~ (Table 2.) were consistently higher in free seedlings than in containers (both P < 0.001). The interaction between photosynthetic capacity, leaf starch, and leaf N on a mass basis was marginally significant (P = 0.058) but A~max~ was highly correlated to both leaf N content and leaf starch (both P < 0.001). Across all measurement campaign A~max~ was higher when foliar N was also higher, usually associated with low levels of leaf starch (Figure `r achemfig`a). A~max~  was also lower when leaf starch was high as higher leaf N often did not coincide with high leaf starch (Figure `r achemfig`b)

\
The photosynthetic parameters J~max~ and Vc~max~ were not different within treatments at the beginning and end of gas exchange campaigns, therefore the parameter means per treatment are reported here (Table 2).  Overall, both J~max~ and Vc~max~ were significantly higher in free seedlings with little variation between soil volume treatments (P = 0.001 & 0.002, respectively). Leaf dark respiration rates were not significantly different across soil volumes (Table 2). The g~1~ parameter, generated for each seedling from the Medlyn et al [-@medlyn2012reconciling] optimal stomatal conductance model, was lowest in the free seedling and was marginally different across soil volume treatments (Table 2). Predicted values of g~s~, using the g~1~ parameter, where highly correlated with observed values (R^2^ = .74, P < 0.001, data not shown).

\
Neither $\Psi$~pd~ nor $\Psi$~l~ were different across treatments, with mean values of -0.27 and -1.2 mPa across all seedlings, respectively. Although g~s~ in free seedlings was generally higher than those in containers (Table 2. P=0.001), the mean rates for all seedlings were high at 0.37 mol H~2~0 m^-2^ s^-1^ and did not decline significantly across the experiment duration. Additionally, leaf $\delta$^13^C at final harvest was not different across treatments (Table 1). Combined these indices provide strong evidence that water stress was not apparent on these well-watered seedlings throughout the experiment. Soil N % at harvest was not different across soil volumes ($\bar{x}$ = 04.5 %) and decreased approximately 3 % across all containers over the experiment duration. This indicates that nutrient leaching from free seedlings or from draining of containers following natural rainfall events did not differ between treatments.

## Modelling seedling biomass
Model M~0~, was able to converge on an optimum LMF (21.6 %) which predicted the mean harvest total biomass of free seedlings within 1.2 %. Using this optimized LMF, the total biomass of modeled seedlings with soil volume restriction (M~1~) was on average 23&pm;2.4 g C more than measured seedlings when comparing against predicted total net leaf C gain (Figure `r massmodelfig`a). Thus, seedling C mass was overestimated by an average of 50&pm;8.7 % in modeled seedlings across the reductions in daily C assimilation per treatment (Figure `r massmodelfig`b). As a result, the reductions in leaf A were not sufficient enough to explain the reduction in harvested seedling biomass with soil volume restriction. This resulted in an unexplained pool of excess C, generated from A, which was not allocated to biomass.

\
The remaining model simulations tested possible C allocation scenarios to account for this excess pool of C. Constraining the model by treatment-specific LMF from the final harvest (M~2~) still overestimated seedling total C by 32&pm;11.1 %, but provided the most improved model predictions (Figure `r massmodelfig2`a). Using harvested LMF, however, does not capture the increase in senescence of seedlings in small containers (Figure `r allomfig`c). Thus, the 7-61 % reduction in measured harvested leaf mass fraction from large to small soil volumes compared to the optimized LMF (M~0~) represents an underestimation of realized leaf C allocation. Increases of 50 % in non-leaf tissue respiration (M~3~) improved biomass estimates slightly but overestimated mass C by an average of 46&pm;9.3 % in seedlings with soil volume restriction (Figure `r massmodelfig2`b). It was further determined that non-tissue respiration rates in modeled seedlings would need to increase by ca. 250 % to account for this entire pool of C.

@@

\
With high rates of g~s~, non-limiting leaf water potential and consistent leaf $\delta$^13^C across soil volume treatments there was little evidence of water stress inducing a limitation to A. This finding is consistent with other container size studies without drought treatments.  For example, reduced A~max~ in cotton seedlings grown at elevated CO~2~ was attributed to sink-limited feedback inhibition from inadequate rooting volume, not decreased g~s~ [@thomas1991root]. Additionally, severe reductions in A in coffee plants was not attributed to impacts of container size on leaf water potentials or g~s~ [@ronchi2006growth]. It is likely that reductions in A of these well watered seedlings was instead the result of limiting soil nutrients on belowground sink strength.

\
Although the soil N pool declined evenly across all treatments leaf N was lowest in the smallest containers, suggesting sink limitation was the greatest in these containers. This makes sense as small containers may reduce N uptake, either from physical root restriction or decreased supply, which affects growth, Rubisco limitation, sugar metabolism, and carbohydrate partitioning between source and sink tissues [@stitt1991rising; @hermans2006plants]. Mycorrhizal colonization could also have been affected in containers. Whether unrestricted mycorrhizal recruitment facilitated the increase in leaf N uptake in free seedlings is unknown. Regardless of whether sink limitation occurred from root or mycorrhizal restriction, the stark contrast in leaf N between free seedlings and seedlings in containers was significantly correlated to reductions in A.

\
As both rooting space and resources were finite in containers, the inability of seedlings to increase the capacity of the belowground C sink resulted in the buildup of C assimilate in leaves. The feedback inhibition of A from starch accumulation has been proposed, yet it is still not known whether there is a starch threshold that triggers the down-regulation process (Nebauer et al. 2011).  Here, declines in A~max~ were correlated with higher starch content throughout the experiment.  This agrees with findings from Equiza et a. (2006) where photosynthetic downregulation from reduced sink strength was more correlated with starch than sugar content in a deciduous conifer.  As starch content in the smallest containers was nearly double that of free seedlings this suggests the response of A to sink inhibition was regulated by this accumulation, as hypothesized.

@@

\
As partitioning to fine roots did not change this provides evidence against an optimal foraging strategy for seedlings in containers. This could be because lateral root development is affected by inanimate obstacles and avoiding growth towards container walls could improve the efficiency of resource allocation [@falik2005root]. Also, the sensitivity of roots to their own exudates near obstructions may be used to adjust growth before stressful conditions occur [@semchenko2008foraging]. Alternatively, physical restriction of root proliferation could have impacted root development and morphology prior to shifts in mass partitioning. Here, increases in root length where detected in several of the soil volume treatments . This is not surprising as plants in containers have been shown to have different root morphology to field grown plants as roots essentially compete with themselves for nutrients [@nesmith1998effect]. Poor soil quality and root restriction, however, likely decreased the capacity of this morphological response to increase N uptake.

@@

\
Increasing rates of non-leaf respiration (M~3~) improved biomass predictions but to a far lesser extent than changes to leaf C allocation. This is still noteworthy as it shows that sink limitation could differentially affect respiration of component tissues. Additionally, keeping rates of respiration fixed in growth models may underestimate impacts of environmental change on tissue respiration above and belowground. The fraction of photosynthate used in respiration varies substantially among species and environments  and is sensitive to changes in growth rates [@Lambers2008]. As C balance is a delicate equilibrium between fluxes of A and respiration, partial accounting of C dynamics can easily lead to erroneous conclusions (Valentini et al. 2000).We agree with Delucia et al. [-@delucia2007forest] that it is likely inappropriate to assume that respiration is a constant fraction of gross primary production in models. Our findings reveal that a combination of different mechanisms, beyond A, is likely at play in driving the observed seedling biomass response to sink manipulation. However, the degree to which these mechanisms will regulate a growth will undoubtedly shift across different experimental manipulations and plant species.

## Conclusions
It has long been known that conditions which affect the photosynthetic process affect growth but that there is also an interrelationship between growth and A [@sweet1966role]. Here, the fate of assimilated C available for plant growth varied between naturally sown seedlings and seedlings with belowground sink inhibition. First, this is important as manipulations of plants grown in containers are often used to draw conclusions about growth and physiological principles but how these results actually reflect field-grown plants has seldom been studied. Second, our finding add more evidence that A and growth are not always entirely synced, an important distinction often missed in studies that manipulate source/sink activity. K&#246rner [-@korner2013growth] suggests that it is the norm for sink activity to feedback onto source activity, causing growth to control A through the demand for C. Although this may be true, we argue that attempts to quantify or at least predict the fate of assimilated C into known pools of growth, storage and C loss are needed prior to addressing this debate. Our modelling results agree with conclusions from Valentine and M&#228;kel&#228; [-@valentine2005bridging] where the problem with predicting tree growth is a problem in forecasting the assimilation and allocation of C and other constituents.  A lack of knowledge regarding C allocation restricts our ability to achieve mass balance and is a major obstacle in understanding the coordination between A and growth. The approach used here has the flexibility to account for multiple drivers of C allocation and provides an avenue to address future questions regarding the impact of environmental change on plant growth.

@@

**Table 1**. Responses of plant and leaf characteristics of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Seedling mass, SRL, root nitrogen and leaf $\delta$^13C values are from final harvest. Values of leaf starch, sugars, nitrogen and SLA represent overall means across measurement campaigns (n=6). Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.

\
**Table 2**. Responses of leaf level gas exchange parameters of *Eucalyptus tereticornis* seedlings to soil volume treatments. Each value reflects the mean(standard error) for each treatment. Units for A~max~ and R~dark~ are $\mu$mol m^-2^ s^-1^ and g~s~ are mol m^-2^ s^-1^, each at 25 &deg;C. Values of A~max~, g~s~ and g~1~ represent overall means across measurement campaigns (n = 6). R~dark~, J~max~ and Vc~max~ values are means of two measurement campaigns at beginning and end of gas exchange measurements. Different letters represent significant differences between treatments. The container effect P value represents the overall difference between seedlings with soil volume restriction and the control seedlings.
